\sc Estimating Evolutionary Parameters for Low Complexity Regions using an Approximate Bayesian Computation

### 2023-01-19 thesis_template/paper0904.tex
Jan 19 Mid-year Updates, almost done

A custom \texttt{C++} script was written to iterate through the algorithm for a desired number of simulations. The Normal distribution was used to control how new parameter values were proposed. We simulated a dataset under the newly proposed parameter value, which consisted of a randomly generated protein sequence. The simulated protein was compared to a protein of known low complexity called SRP40, which is found in the model organism \sclong. The protein sequence for SRP40 was downloaded from the NCBI database. To quantitatively compare similarities between simulated and observed protein sequences, vectors of summary statistics (characteristics that describe the sequences) were created for each sequence. The euclidean distance between the two vectors was calculated and compared to a threshold value to determine if the newly proposed parameter value could be accepted.

@@

\subsection{Simulation Step: Creation and Mutation of Protein Sequences}
% Bulk of methods, talk about the simulation step which involves creating proteins
% mutating proteins (indels and point mutations) and how we scan for repeats and assign deviates. the whole process, even generating deviates here

To determine

\subsection{Euclidean Distance Calculation}
% May be important to highlight this calculation because we use it to normalize the vectors of summary statistics before examining results.
% Could make this a weighted Euclidean Distance?
% Can I also put normalizing the vectors here?

To determine the similarity between simulated and observed protein sequences, the distance between the vectors of summary statistics was calculated by employing a Euclidean distance measure \eqref{eq:1}.

\begin{equation}
	d(P1,P2) = \sqrt{(x_2-x_1)^2 + (y_2-y_1)^2 + (z_2-z_1)^2 }
	\label{eq:1}
\end{equation}

>>>>

A custom \texttt{C++} script was written to iterate through the algorithm for a desired number of simulations. The Normal distribution was used to control how new parameter values were proposed \hl{SHOULD I MENTION THE MEAN AND STDEV OF THE NORMAL HERE? WHAT IF I WANT TO TEST MANY DIFFERENT KINDS}. We simulated a dataset under the newly proposed parameter value, which consisted of a randomly generated protein sequence. The simulated protein was compared to a protein of known low complexity called SRP40, which is found in the model organism \sclong. The protein sequence for SRP40 was downloaded from the NCBI database. To quantitatively compare similarities between simulated and observed protein sequences, vectors of summary statistics (characteristics that describe the sequences) were created for each sequence. The euclidean distance between the two vectors was calculated and compared to a threshold value to determine if the newly proposed parameter value could be accepted.

@@

\hl{Need Help:} We currently do not know how well these summary statistics explain the data. Weighting summary statistics could be beneficial, but determining how we weight each one appropriately is challenging at the moment. There are potential weighting techniques, such as one method used by \citet{hamilton2005bayesian} in which summary statistics are regressed on retained parameters, allowing for a weight to be calculated for each summary statistic and subsequently used when calculating the Euclidean Distance. This method is very difficult to implement and I am struggling. Maybe other alternatives for this weighting?

\subsection{Simulation Step: Creation and Mutation of Protein Sequences}
% Bulk of methods, talk about the simulation step which involves creating proteins
% mutating proteins (indels and point mutations) and how we scan for repeats and assign deviates. the whole process, even generating deviates here
%Scale parameter is beta for exponential distribution

To bypass calculation of the likelihood function, a custom \texttt{C++} script was written to simulate the generation and mutation of protein sequences over numerous generations. A random protein of desired length was generated first. We generated a protein sequence 400 amino acids in length, similar to the length of the SRP40 protein in \scshrt. This simulated protein sequence was then  mutated in the following ways.

We iterated over the simulated protein sequence and assigned exponential deviates to each amino acid. We utilized the exponential distribution because it is commonly used to model waiting times between events. Mutation and indel rates served to act as the scale parameter ($\beta$), or mean of the distribution, indicative of the mean time until mutation occurs. In the case of mutation rate, the same rate was utilized across all sites, with the assumption made that repeats do not play a role in point mutation. In the case of the indel rate, we scanned for repeats, and if found, we multiplied the length of the repetitive segment by the indel rate, and utilized this new value as the scale parameter ($\beta$) for the exponential distribution. \hl{THINK ABOUT WHAT THIS MEANS IN TERMS OF THE EXPONENTIAL DISTRIBUTION, THE LONGER THE LENGTH OF REPEAT, HOW DOES THIS AFFECT THE EXPONENTIAL DISTRIBUTION, DOES IT MAKE SENSE?}.

Exponential deviates were stored in two vectors, one for deviates generated using mutation rate, and the other for deviates generated using the indel rate. We then identified a single deviate with the lowest value (based on both vectors), which represented the residue that mutated quickest. Depending on which vector the lowest deviate came from, we either altered the corresponding amino acid or inserted/deleted a repetitive amino acid.

We ran the simulation ten times for each newly proposed parameter value

\subsection{Normalization and Euclidean Distance Calculation}
% May be important to highlight this calculation because we use it to normalize the vectors of summary statistics before examining results.
% Could make this a weighted Euclidean Distance?
% Can I also put normalizing the vectors here?

To determine the similarity between simulated and observed protein sequences, the distance between the vectors of summary statistics was calculated by employing a Euclidean distance measure \eqref{eq:2}. Before calculating this distance, vectors of summary statistics were normalized \eqref{eq:1} to prevent large values (such as length) from dominating the distance measure.

\begin{equation}
	length = \sqrt{(x*x) + (y*y) + (z*z)} \\
	\label{eq:1}
\end{equation}

\begin{center}
	$normalized = [x/length, y/length, z/length]$
\end{center}

\begin{equation}
	d(P1,P2) = \sqrt{(x_2-x_1)^2 + (y_2-y_1)^2 + (z_2-z_1)^2 }
	\label{eq:2}
\end{equation}

\hl{Need Help:}
